Optimizing and Incrementalizing Collection Queries by AST Transformations

### 2017-07-27 chap-intro.tex
Intro revisions

What's worse, some of these manual optimizations are best done over the whole
program. Some optimizations only become possible after inlining, which breaks
modularity if done by hand.
Worse, adding an index on some collection can speed up looking
for some information, but each index must be maintained incrementally when the
underlying data changes. Depending on the actual queries and updates performed
on the collection, and on how often they happen, it might turn out that updating
an index takes more time than the index saves; hence the choice of which indexes
to enable depends on the whole program. However, adding/removing an index
requires updating all PL queries to use it, while RDBMS queries can use an index
transparently.

@@

The first chapters of \cref{part:incr} are originally based on work by \citet*{CaiEtAl2014ILC}, though
significantly revised; only \cref{sec:applying,sec:rw} survive mostly unchanged.
This work was even more of a team effort. I initiated and led the
overall project and came up with the original notion of change structures; Cai
Yufei contributed differentiation itself and its first correctness proof.
Tillmann Rendel contributed significant simplifications and started the overall
mechanization effort.
\Cref{part:incr} contains novel correctness proof; its history and
contributions are discussed in \cref{sec:ilc-dev-history}.

Furthermore, \cref{ch:cts} comes from a recently submitted
manuscript~\citep*{Giarrusso2018Static}. I
designed the overall approach, the transformation and the case study on
sequences and nested loops. Proofs were done in collaboration with Yann
Régis-Gianas: most of the actual mechanization work is his, though I contributed significantly to the
correctness proof for ILC\@. Philipp Schuster contributed to the evaluation,
devising language plugins for bags and maps in collaboration with me.

>>>>

What's worse, some of these manual optimizations are best done over the whole
program. Some optimizations only become possible after inlining, which reduces
modularity if done by hand.\footnote{%
In \cref{sec:aosd13-intro} we define modularity as the ability to abstract
behavior in a separate function (possibly part of a different module) to enable
reuse and improve understandability.%
}

Worse, adding an index on some collection can speed up looking
for some information, but each index must be maintained incrementally when the
underlying data changes. Depending on the actual queries and updates performed
on the collection, and on how often they happen, it might turn out that updating
an index takes more time than the index saves; hence the choice of which indexes
to enable depends on the whole program. However, adding/removing an index
requires updating all PL queries to use it, while RDBMS queries can use an index
transparently.

@@

The first chapters of \cref{part:incr} are originally based on work by \citet*{CaiEtAl2014ILC}, though
significantly revised; only \cref{sec:applying,sec:rw} survive mostly unchanged.
This work was even more of a team effort. I initiated and led the
overall project and came up with the original notion of change structures. Cai
Yufei contributed differentiation itself and its first correctness proof.
Tillmann Rendel contributed significant simplifications and started the overall
mechanization effort. Klaus Ostermann provided senior supervision that proved
essential to the project.
\Cref{part:incr} contains a novel correctness proof for simply-typed $\lambda$-calculus;
for its history and contributions \cref{sec:ilc-dev-history}.

Furthermore, \cref{ch:cts} comes from a recently submitted
manuscript~\citep*{Giarrusso2018Static}. I
designed the overall approach, the transformation and the case study on
sequences and nested loops. Proofs were done in collaboration with Yann
Régis-Gianas: he is the main author of the Coq mechanization and proof, though I
contributed significantly to the correctness proof for ILC\@, in particular with
the proofs described in \cref{ch:bsos} and their partial Agda mechanization.
The history of this correctness proof is described in
\cref{sec:ilc-bsos-dev-history}.
Philipp Schuster contributed to the evaluation,
devising language plugins for bags and maps in collaboration with me.
